---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version-ago/mapsapi)](https://cran.r-project.org/package=mapsapi)
[![CRAN_Downloads_Badge](http://cranlogs.r-pkg.org/badges/last-month/mapsapi)](https://cran.r-project.org/package=mapsapi)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# mapsapi

The `mapsapi` package provides an interface to the Google Maps APIs, currently four of them:

* <a href="https://developers.google.com/maps/documentation/directions/" target="_blank">Google Maps Direction API</a>
* <a href="https://developers.google.com/maps/documentation/distance-matrix/" target="_blank">Google Maps Distance Matrix API</a>
* <a href="https://developers.google.com/maps/documentation/geocoding/" target="_blank">Google Maps Geocode API</a>
* <a href="https://developers.google.com/maps/documentation/maps-static/" target="_blank">Maps Static API</a>

Functions `mp_directions`, `mp_matrix` and `mp_geocode` are used to access the Directions, Matrix and Geocode APIs, respectively. They return an `xml_document` object (package `xml2`) with the response contents.  

* Given a *directions* response, functions `mp_get_routes` and `mp_get_segments` can be used to process the response document into a spatial layer. Function `mp_get_routes` gives each alternative as a separate line, while function `mp_get_segments` gives each segment (that is, a portion of the route associated with specific driving instructions) as a separate line. 

* Given a *distance matrix* response, function `mp_get_matrix` can be used to obtain distance/duration matrices.

* Given a *geocode* response, functions `mp_get_points` and `mp_get_bounds` can be used to obtain geocoded locations as a point or polygon (bounds) layer. 

The fourth function `mp_map` is used to access the Maps Static API. It returns a `stars` raster RGB image, which can be used as background in maps.

## Installation

The CRAN version can be installed with:

```{r, eval=FALSE}
install.packages("mapsapi")
```

The development version can be installed using `remotes`:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("michaeldorman/mapsapi")
```

Once installed, the package can be loaded with `library`:

```{r}
library(mapsapi)
```

**Note: due to [new Google Maps API policy](https://developers.google.com/maps/billing/important-updates), starting from June 2018 the functions require an API key.**

```{r, eval=FALSE}
key = "AIz....."
```

## Documentation

The complete documentation can be found at [https://michaeldorman.github.io/mapsapi/](https://michaeldorman.github.io/mapsapi/).

## Example

The following code section obtains (and plots) the driving directions from New-York to Los Angeles. 

```{r, include=FALSE}
key = readLines("~/key")
```

```{r example}
# Get routes (XML document)
doc = mp_directions(
  origin = "New-York",
  destination = "Los Angeles",
  alternatives = TRUE,
  key = key, 
  quiet = TRUE
)

# Extract lines 'sf' layer
r = mp_get_routes(doc)

# Plot
library(maps)
library(sf)
map("state", fill = FALSE, col = "grey")
plot(st_geometry(r), col = c("red", "green", "blue"), add = TRUE)
```

```{r, include=FALSE}
x = list.files(pattern = "^README-.*\\.png$")
file.copy(x, "docs/", overwrite = TRUE)
```
